---
title: "p8105_hw2_ab4669"
author: "Adrian Blanco"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install packages

```{r setup, include=FALSE}

install.packages("readr")
install.packages("janitor")

```



# Libraries

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
library(readr)
library(janitor)

```


# Problem 1
This problem focuses on NYC Transit data; in particular, this CSV file contains information related to each entrance and exit for each subway station in NYC.

#### Read the data

```{r setup, include=FALSE}

nyc_transit_data <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

```

#### Clean the data

```{r setup, include=FALSE}

nyc_transit_data_clean <- nyc_transit_data %>% 
  janitor::clean_names()

```

#### Select the data

```{r setup, include=FALSE}

nyc_transit_data_clean <- nyc_transit_data_clean %>%  
  select(line:entry, 
         vending, 
         ada)

```

#### Convert entry from character to logical

```{r setup, include=FALSE}

nyc_transit_data_clean$entry <- ifelse(nyc_transit_data_clean$entry == "YES", TRUE, FALSE)

```





Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? 
Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

```{r setup, include=FALSE}

nrow(distinct(nyc_transit_data_clean, line, station_name))

```

465

How many stations are ADA compliant?

```{r setup, include=FALSE}

nrow(filter(distinct(nyc_transit_data_clean, line, station_name, ada), ada == TRUE))

```

84

What proportion of station entrances / exits without vending allow entrance?

```{r setup, include=FALSE}

nrow(filter(nyc_transit_data_clean, vending == "NO" & entry == TRUE))/nrow(nyc_transit_data_clean)

```

Reformat data so that route number and route name are distinct variables.

```{r setup, include=FALSE}

nyc_transit_data_reformatted <- nyc_transit_data_clean %>%
                      gather(key = route_number, value = route_name, route1:route11) %>%
                      separate(route_number,
                               into = c("delete", "route_number"),
                               sep = 5) %>%
                      select(-delete)

```

How many distinct stations serve the A train? 

```{r setup, include=FALSE}

nrow(filter(distinct(nyc_transit_data_reformatted, line, station_name, route_name), route_name == "A"))

```



Of the stations that serve the A train, how many are ADA compliant?

```{r setup, include=FALSE}

nrow(filter(distinct(nyc_transit_data_reformatted, line, station_name, route_name, ada), route_name == "A", ada == TRUE))

```

# Problem 2
This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website. Please use the  HealthyHarborWaterWheelTotals2017-9-26.xlsx version.

Read and clean the Mr. Trash Wheel sheet:

#### Read the data

```{r setup, include=FALSE}

trash_wheel_data <- read_excel("./data/HealthyHarborWaterWheelTotals2017-3.xlsx", 
                               sheet = "Mr. Trash Wheel", 
                               range = cell_cols("A:N")) 

```



specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
use reasonable variable names
omit rows that do not include dumpster-specific data
rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)
Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?

# Problem 3
This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package.

#### Installation

```{r setup, include=FALSE}

devtools::install_github("p8105/p8105.datasets")

```

```{r setup, include=FALSE}

library(p8105.datasets)

```


#### Read the data

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010

```



For this question:

format the data to use appropriate variable names;

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names()

```

focus on the “Overall Health” topic

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health")

```

exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == 'Overall Health') %>%
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))

```

structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == 'Overall Health') %>%
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names()

```

create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

```{r setup, include=FALSE}

BRFSS_data <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == 'Overall Health') %>%
  select(-class, -topic, -question, -sample_size, -c(confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>%
  mutate(excel_or_verygood = excellent + very_good)

```

Using this dataset, do or answer the following:

How many unique locations are included in the dataset? 

```{r setup, include=FALSE}

unique_locations <- BRFSS_data %>%
  distinct(locationabbr) %>% 
  count()

unique_locations

```

51 unique_locations

Is every state represented? 


What state is observed the most?

In 2002, what is the median of the “Excellent” response value?

```{r setup, include=FALSE}

excellent2002 <- BRFSS_data %>% 
  filter(year == 2002) %>%
  select(excellent) 

median(excellent2002$excellent, na.rm = TRUE)

```

Median: 23.6


Make a histogram of “Excellent” response values in the year 2002.

```{r setup, include=FALSE}

hist(excellent2002$excellent, main = "Histogram of Excellent response values in 2002.", xlab = "excellent values" )

```


Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.

```{r setup, include=FALSE}

BRFSS_data %>% 
  filter(locationdesc == "NY - New York County" | locationdesc == "NY - Queens County") %>%
  mutate(excellent_proportion = excellent/(excellent + fair + poor + good + very_good)) %>% 
  ggplot(aes(x = excellent_proportion, y = year)) + geom_point(aes(color = locationdesc))

```